---
title: "Case Study 02 DDS"
date: "11/20/2019"
output: html_document
---

#Case Study 02 - Final Project

Youtube Video Presenation at: https://youtu.be/g9Bf-9vVnF4

CaseStudy 2 Github Repository website: https://github.com/adebouse/Case-Study-2.git

Github website: adebouse.github.io


Introduction:

As a representative of DDS Analytics, I will focus my analysis on identifying varaibles a company can focus on to improve employee development and increase retention rates. Analyzing the frequencies of categorical variables and the medians of continuous variables for specific job titles we hope to identify key predictors that identify employees who are likely to remain and those who will leave/be fired. 

Summary:

We discovered that Overtime, Stock Option and Job Level had a strong correlation with predicting Attrition, and it was possible that those job position with turnover rates higher than the overall company could be due high levels of retirement. 



```{r}
library(readr)
library(tidyverse)
library(GGally)
library(corrplot)
library(DataExplorer)
library(e1071)
library(caret)
library(class)
library(GoodmanKruskal)
library(PerformanceAnalytics)
library(scales)

 

#Import CaseStudy2 dataset

CaseStudy2 <- read.csv("C:/Users/Antonio Debouse/Desktop/Doing Data Science/GitHub/Case-Study-2/CaseStudy2-data.csv")
 

#Check for missing data

sum(is.na(CaseStudy2))

 

##No missing data

 

#Inspect dataset variables

str(CaseStudy2)

head(CaseStudy2)

#Change varibles to factors

names <- c('Education','EnvironmentSatisfaction','JobInvolvement','JobLevel', 'JobSatisfaction','PerformanceRating','RelationshipSatisfaction','StockOptionLevel','WorkLifeBalance')

 
#Change character variables to factors
CaseStudy2[,names] <- lapply(CaseStudy2[,names] , factor)

str(CaseStudy2)

#Reorder columns with factors and integers grouped together, eliminate unneeded columns

CaseStudy2_reorg <- CaseStudy2[,c(4,6,8,9,12,13,15:19,24,26,27,29,32,2,5,7,14,20,21,22,25,30,31,33:36,3)]

view(CaseStudy2_reorg)

str(CaseStudy2_reorg)

 
 
```


```{r}


#plot Overall Attrition frequency


CaseStudy2_reorg %>% ggplot(aes(x=Attrition, fill =Attrition))+
  geom_bar(aes(y = (..count..)/sum(..count..)))+
  geom_text(aes(y = ((..count..)/sum(..count..)), label = scales::percent((..count..)/sum(..count..))), stat = "count", vjust = -0.25) +
  scale_y_continuous(labels = percent) +
  ggtitle("Frequency of Attrition Chart")+
  ylab("Percentage")


```

The company has retained only 16% of employees with 84% having exited the company; therefore, my target is to create a model with an accuracy of greater than 84% since that is the accuracy I would get by guessing "no" everytime.

 

 

```{r}

#Identify any trends specific to any job role

##Sum of Observations for each Job Role
print(CaseStudy2_reorg %>%
  group_by(JobRole) %>%
  summarize(count= n()) %>%
  arrange(desc(count)))

  

##Barchart(count) for job roles

CaseStudy2_reorg %>% ggplot(aes(x= JobRole, fill = JobRole)) +
  geom_bar() +
  coord_flip()+
  geom_text(aes(label=..count..),stat = "count",hjust = -0.25)+
  ggtitle("Chart of Job Roles Sums")+
 theme(legend.position = "none")

 

##Barchart(percentage) for job roles

CaseStudy2_reorg %>% ggplot(aes(x=JobRole, fill =JobRole))+
  geom_bar(aes(y = (..count..)/sum(..count..)))+
  geom_text(aes(y = ((..count..)/sum(..count..)), label = scales::percent((..count..)/sum(..count..))), stat = "count", hjust = -0.25) +
  scale_y_continuous(labels = percent) +
  ggtitle("JobRole Percentage Breakout Chart")+
  ylab("Percentage")+
  theme(legend.position = "none")+
  coord_flip()

##Create bar chart with Attrition percentages grouped by JobRole
CaseStudy2_reorg %>% ggplot(aes(x=JobRole, fill =Attrition))+
  geom_bar(aes(y = (..count..)/sum(..count..)))+
  geom_text(aes(y = ((..count..)/sum(..count..)), label = scales::percent((..count..)/sum(..count..)), vjust = -0.25), stat = "count") +
  scale_y_continuous(labels = percent) +
  ggtitle("Frequency of Attrition Chart")+
  ylab("Percentage")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 

##Create Frequency table for Attrition grouped by JobRole
JobRoleAttrition_Frequency <- table(CaseStudy2_reorg$JobRole,CaseStudy2_reorg$Attrition)

##Feed JobRoleAttrition frequency table to prop.table() to produce relative percentages
JobRoleAttrition_Percentages <-round(prop.table(JobRoleAttrition_Frequency,margin = 1),2)

#Place table into dataframe
dataframecovert <- as.data.frame(JobRoleAttrition_Percentages)
dataframecovert$Freq <- dataframecovert$Freq * 100

#Change column names
colnames(dataframecovert) <- c("Job Title", "Attrition","Freq")


view(dataframecovert)

#Statistical Summary of Attrition = No, group by JobRole
dataframecovert %>% filter(Attrition == "No") %>%
  summary(dataframecovert$Freq) 

#Standard deviation of Attrition = No, group by JobRole
dataframecovert %>% filter(Attrition == "No") %>%
  mutate(standev = sd(Freq)) 


#Create a column chart with relative percentage of Attrition grouped by JobRole

dataframecovert %>% ggplot(aes(x=`Job Title`, fill =Attrition,y=Freq))+
  geom_col()+
  geom_text(aes(label=Freq,),position = position_stack(vjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Percentage of Attrition Grouped by Job Tiltle")+
  ylab("Percentage")



###Create Frequency table for EducationField, Grouped by JobRole and sub-grouped by Attrition (yes/no)
###ftable(CaseStudy2_reorg$JobRole,CaseStudy2_reorg$Attrition,CaseStudy2_reorg$EducationField)




```



```{r} 
##Create Frequency table for BusinessTravel grouped by JobRole
JobRole_BusinessTravel_Frequency <- table(CaseStudy2_reorg$JobRole,CaseStudy2_reorg$BusinessTravel)

##Feed JobRoleBusinessTravel frequency table to prop.table() to produce relative percentages
JobRole_BusinessTravel_Percentages <-round(prop.table(JobRole_BusinessTravel_Frequency,margin = 1),2)

#Place table into dataframe
dataframecovert <- as.data.frame(JobRole_BusinessTravel_Percentages)
dataframecovert$Freq <- dataframecovert$Freq * 100

#Change column names
colnames(dataframecovert) <- c("Job Title", "BusinessTravel","Freq")


view(dataframecovert)


#Create a column chart with relative percentage of BusinessTravel grouped by JobRole

dataframecovert %>% ggplot(aes(x=`Job Title`, fill =BusinessTravel,y=Freq))+
  geom_col()+
  geom_text(aes(label=Freq,),position = position_stack(vjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Percentage of BusinessTravel Grouped by Job Tiltle")+
  ylab("Percentage")
 

```

 
```{r}
##Create Frequency table for Education grouped by JobRole
JobRole_Education_Frequency <- table(CaseStudy2_reorg$JobRole,CaseStudy2_reorg$Education)

##Feed JobRole_Education frequency table to prop.table() to produce relative percentages
JobRole_Education_Percentages <-round(prop.table(JobRole_Education_Frequency,margin = 1),2)

#Place table into dataframe
dataframecovert <- as.data.frame(JobRole_Education_Percentages)
dataframecovert$Freq <- dataframecovert$Freq * 100

#Change column names
colnames(dataframecovert) <- c("Job Title", "Education","Freq")


view(dataframecovert)


#Create a column chart with relative percentage of Education grouped by JobRole

dataframecovert %>% ggplot(aes(x=`Job Title`, fill =Education,y=Freq))+
  geom_col()+
  geom_text(aes(label=Freq,),position = position_stack(vjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Percentage of Education Grouped by Job Tiltle")+
  ylab("Percentage")


```
 
```{r}
##Create Frequency table for EducationField grouped by JobRole
JobRole_EducationField_Frequency <- table(CaseStudy2_reorg$JobRole,CaseStudy2_reorg$EducationField)

##Feed JobRole_EducationField frequency table to prop.table() to produce relative percentages
JobRole_EducationField_Percentages <-round(prop.table(JobRole_EducationField_Frequency,margin = 1),2)

#Place table into dataframe
dataframecovert <- as.data.frame(JobRole_EducationField_Percentages)
dataframecovert$Freq <- dataframecovert$Freq * 100

#Change column names
colnames(dataframecovert) <- c("Job Title", "EducationField","Freq")


#Create a column chart with relative percentage of EducationField grouped by JobRole

dataframecovert %>% ggplot(aes(x=`Job Title`, fill =EducationField,y=Freq))+
  geom_col()+
  geom_text(aes(label=Freq,),position = position_stack(vjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Percentage of EducationField Grouped by Job Tiltle")+
  ylab("Percentage")


```


```{r}
##Create Frequency table for EnvironSatisfaction grouped by JobRole
JobRole_EnvironSatisfaction_Frequency <- table(CaseStudy2_reorg$JobRole,CaseStudy2_reorg$EnvironmentSatisfaction)

##Feed JobRole_EnvironSatisfaction frequency table to prop.table() to produce relative percentages
JobRole_EnvironSatisfaction_Percentages <-round(prop.table(JobRole_EnvironSatisfaction_Frequency,margin = 1),2)

#Place table into dataframe
dataframecovert <- as.data.frame(JobRole_EnvironSatisfaction_Percentages)
dataframecovert$Freq <- dataframecovert$Freq * 100

#Change column names
colnames(dataframecovert) <- c("Job Title", "EnvironSatisfaction","Freq")



#Create a column chart with relative percentage of EnvironSatisfaction grouped by JobRole

dataframecovert %>% ggplot(aes(x=`Job Title`, fill =EnvironSatisfaction,y=Freq))+
  geom_col()+
  geom_text(aes(label=Freq,),position = position_stack(vjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Percentage of EnvironSatisfaction Grouped by Job Tiltle")+
  ylab("Percentage")


```



```{r}
##Create Frequency table for JobInvolvement grouped by JobRole
JobRole_JobInvolvement_Frequency <- table(CaseStudy2_reorg$JobRole,CaseStudy2_reorg$JobInvolvement)

##Feed JobRole_JobInvolvement frequency table to prop.table() to produce relative percentages
JobRole_JobInvolvement_Percentages <-round(prop.table(JobRole_JobInvolvement_Frequency,margin = 1),2)

#Place table into dataframe
dataframecovert <- as.data.frame(JobRole_JobInvolvement_Percentages)
dataframecovert$Freq <- dataframecovert$Freq * 100

#Change column names
colnames(dataframecovert) <- c("Job Title", "JobInvolvement","Freq")



#Create a column chart with relative percentage of JobInvolvement grouped by JobRole

dataframecovert %>% ggplot(aes(x=`Job Title`, fill =JobInvolvement,y=Freq))+
  geom_col()+
  geom_text(aes(label=Freq,),position = position_stack(vjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Percentage of JobInvolvement Grouped by Job Tiltle")+
  ylab("Percentage")


```
```{r}
##Create Frequency table for MaritalStatus grouped by JobRole
JobRole_MaritalStatus_Frequency <- table(CaseStudy2_reorg$JobRole,CaseStudy2_reorg$MaritalStatus)

##Feed JobRole_MaritalStatus frequency table to prop.table() to produce relative percentages
JobRole_MaritalStatus_Percentages <-round(prop.table(JobRole_MaritalStatus_Frequency,margin = 1),2)

#Place table into dataframe
dataframecovert <- as.data.frame(JobRole_MaritalStatus_Percentages)
dataframecovert$Freq <- dataframecovert$Freq * 100

#Change column names
colnames(dataframecovert) <- c("Job Title", "MaritalStatus","Freq")



#Create a column chart with relative percentage of MaritalStatus grouped by JobRole

dataframecovert %>% ggplot(aes(x=`Job Title`, fill =MaritalStatus,y=Freq))+
  geom_col()+
  geom_text(aes(label=Freq,),position = position_stack(vjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Percentage of MaritalStatus Grouped by Job Tiltle")+
  ylab("Percentage")


```
 
```{r}
##Create Frequency table for Overtime grouped by JobRole
JobRole_Overtime_Frequency <- table(CaseStudy2_reorg$JobRole,CaseStudy2_reorg$OverTime)

##Feed JobRole_Overtime frequency table to prop.table() to produce relative percentages
JobRole_Overtime_Percentages <-round(prop.table(JobRole_Overtime_Frequency,margin = 1),2)

#Place table into dataframe
dataframecovert <- as.data.frame(JobRole_Overtime_Percentages)
dataframecovert$Freq <- dataframecovert$Freq * 100

#Change column names
colnames(dataframecovert) <- c("Job Title", "Overtime","Freq")



#Create a column chart with relative percentage of Overtime grouped by JobRole

dataframecovert %>% ggplot(aes(x=`Job Title`, fill =Overtime,y=Freq))+
  geom_col()+
  geom_text(aes(label=Freq,),position = position_stack(vjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Percentage of Overtime Grouped by Job Tiltle")+
  ylab("Percentage")


```

```{r}
##Create Frequency table for PerformanceRating grouped by JobRole
JobRole_PerformanceRating_Frequency <- table(CaseStudy2_reorg$JobRole,CaseStudy2_reorg$PerformanceRating)

##Feed JobRole_PerformanceRating frequency table to prop.table() to produce relative percentages
JobRole_PerformanceRating_Percentages <-round(prop.table(JobRole_PerformanceRating_Frequency,margin = 1),2)

#Place table into dataframe
dataframecovert <- as.data.frame(JobRole_PerformanceRating_Percentages)
dataframecovert$Freq <- dataframecovert$Freq * 100

#Change column names
colnames(dataframecovert) <- c("Job Title", "PerformanceRating","Freq")



#Create a column chart with relative percentage of PerformanceRating grouped by JobRole

dataframecovert %>% ggplot(aes(x=`Job Title`, fill =PerformanceRating,y=Freq))+
  geom_col()+
  geom_text(aes(label=Freq,),position = position_stack(vjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Percentage of PerformanceRating Grouped by Job Tiltle")+
  ylab("Percentage")


```
 
```{r}
##Create Frequency table for RelationshipSatisfaction grouped by JobRole
JobRole_RelationshipSatisfaction_Frequency <- table(CaseStudy2_reorg$JobRole,CaseStudy2_reorg$RelationshipSatisfaction
)

##Feed JobRole_RelationshipSatisfaction frequency table to prop.table() to produce relative percentages
JobRole_RelationshipSatisfaction_Percentages <-round(prop.table(JobRole_RelationshipSatisfaction_Frequency,margin = 1),2)

#Place table into dataframe
dataframecovert <- as.data.frame(JobRole_RelationshipSatisfaction_Percentages)
dataframecovert$Freq <- dataframecovert$Freq * 100

#Change column names
colnames(dataframecovert) <- c("Job Title", "RelationshipSatisfaction","Freq")



#Create a column chart with relative percentage of RelationshipSatisfaction grouped by JobRole

dataframecovert %>% ggplot(aes(x=`Job Title`, fill =RelationshipSatisfaction,y=Freq))+
  geom_col()+
  geom_text(aes(label=Freq,),position = position_stack(vjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Percentage of RelationshipSatisfaction Grouped by Job Tiltle")+
  ylab("Percentage")


```



```{r}
##Create Frequency table for StockOptionLevel grouped by JobRole
JobRole_StockOptionLevel_Frequency <- table(CaseStudy2_reorg$JobRole,CaseStudy2_reorg$StockOptionLevel
)

##Feed JobRole_StockOptionLevel frequency table to prop.table() to produce relative percentages
JobRole_StockOptionLevel_Percentages <-round(prop.table(JobRole_StockOptionLevel_Frequency,margin = 1),2)

#Place table into dataframe
dataframecovert <- as.data.frame(JobRole_StockOptionLevel_Percentages)
dataframecovert$Freq <- dataframecovert$Freq * 100

#Change column names
colnames(dataframecovert) <- c("Job Title", "StockOptionLevel","Freq")



#Create a column chart with relative percentage of StockOptionLevel grouped by JobRole

dataframecovert %>% ggplot(aes(x=`Job Title`, fill =StockOptionLevel,y=Freq))+
  geom_col()+
  geom_text(aes(label=Freq,),position = position_stack(vjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Percentage of StockOptionLevel Grouped by Job Tiltle")+
  ylab("Percentage")


```
```{r}
##Create Frequency table for WorkLifeBalance grouped by JobRole
JobRole_WorkLifeBalance_Frequency <- table(CaseStudy2_reorg$JobRole,CaseStudy2_reorg$WorkLifeBalance
)

##Feed JobRole_WorkLifeBalance frequency table to prop.table() to produce relative percentages
JobRole_WorkLifeBalance_Percentages <-round(prop.table(JobRole_WorkLifeBalance_Frequency,margin = 1),2)

#Place table into dataframe
dataframecovert <- as.data.frame(JobRole_WorkLifeBalance_Percentages)
dataframecovert$Freq <- dataframecovert$Freq * 100

#Change column names
colnames(dataframecovert) <- c("Job Title", "WorkLifeBalance","Freq")



#Create a column chart with relative percentage of WorkLifeBalance grouped by JobRole

dataframecovert %>% ggplot(aes(x=`Job Title`, fill =WorkLifeBalance,y=Freq))+
  geom_col()+
  geom_text(aes(label=Freq,),position = position_stack(vjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Percentage of WorkLifeBalance Grouped by Job Tiltle")+
  ylab("Percentage")


```

```{r}
##Statistical Summary of Continuous Variables grouped by JobRole
##medians of continuous variables grouped by JobRole

MediansofContinousVars <- CaseStudy2_reorg %>% group_by(JobRole) %>%
  summarise(Median_Age = median(Age),
            Median_DailyRate = median(DailyRate),
            Median_DistanceFromHome = median(DistanceFromHome),
            Median_HourlyRate = median(HourlyRate),
            Median_MonthlyIncome = median(MonthlyIncome),
            Median_MonthlyRate = median(MonthlyRate),
            Median_NumCompaniesWorked = median(NumCompaniesWorked),
            Median_PercentSalaryHike = median(PercentSalaryHike),
            Median_TotalWorkingYears = median(TotalWorkingYears),
            Median_TrainTimesLastYear = median(TrainingTimesLastYear),
            Median_YearsAtCompany = median(YearsAtCompany),
            Median_YearsINCurrentRole = median(YearsInCurrentRole),
            Median_YearsSinceLastPromotion = median(YearsSinceLastPromotion),
            Median_YearsWithCurrManager = median(YearsWithCurrManager))
  
summary(MediansofContinousVars)

#Medians for Healthcare Rep
MediansofHCrepVars <- CaseStudy2_reorg %>% group_by(JobRole=="Healthcare Representative") %>%
  summarise(Median_Age = median(Age),
            Median_DailyRate = median(DailyRate),
            Median_DistanceFromHome = median(DistanceFromHome),
            Median_HourlyRate = median(HourlyRate),
            Median_MonthlyIncome = median(MonthlyIncome),
            Median_MonthlyRate = median(MonthlyRate),
            Median_NumCompaniesWorked = median(NumCompaniesWorked),
            Median_PercentSalaryHike = median(PercentSalaryHike),
            Median_TotalWorkingYears = median(TotalWorkingYears),
            Median_TrainTimesLastYear = median(TrainingTimesLastYear),
            Median_YearsAtCompany = median(YearsAtCompany),
            Median_YearsINCurrentRole = median(YearsInCurrentRole),
            Median_YearsSinceLastPromotion = median(YearsSinceLastPromotion),
            Median_YearsWithCurrManager = median(YearsWithCurrManager))
summary(MediansofHCrepVars)


#Medians for Human Resources
MediansofHR <- CaseStudy2_reorg %>% group_by(JobRole=="Human Resources") %>%
  summarise(Median_Age = median(Age),
            Median_DailyRate = median(DailyRate),
            Median_DistanceFromHome = median(DistanceFromHome),
            Median_HourlyRate = median(HourlyRate),
            Median_MonthlyIncome = median(MonthlyIncome),
            Median_MonthlyRate = median(MonthlyRate),
            Median_NumCompaniesWorked = median(NumCompaniesWorked),
            Median_PercentSalaryHike = median(PercentSalaryHike),
            Median_TotalWorkingYears = median(TotalWorkingYears),
            Median_TrainTimesLastYear = median(TrainingTimesLastYear),
            Median_YearsAtCompany = median(YearsAtCompany),
            Median_YearsINCurrentRole = median(YearsInCurrentRole),
            Median_YearsSinceLastPromotion = median(YearsSinceLastPromotion),
            Median_YearsWithCurrManager = median(YearsWithCurrManager))
summary(MediansofHR)

#Medians for Laboratory Technician
MediansofLabtech <- CaseStudy2_reorg %>% group_by(JobRole=="Laboratory Technician") %>%
  summarise(Median_Age = median(Age),
            Median_DailyRate = median(DailyRate),
            Median_DistanceFromHome = median(DistanceFromHome),
            Median_HourlyRate = median(HourlyRate),
            Median_MonthlyIncome = median(MonthlyIncome),
            Median_MonthlyRate = median(MonthlyRate),
            Median_NumCompaniesWorked = median(NumCompaniesWorked),
            Median_PercentSalaryHike = median(PercentSalaryHike),
            Median_TotalWorkingYears = median(TotalWorkingYears),
            Median_TrainTimesLastYear = median(TrainingTimesLastYear),
            Median_YearsAtCompany = median(YearsAtCompany),
            Median_YearsINCurrentRole = median(YearsInCurrentRole),
            Median_YearsSinceLastPromotion = median(YearsSinceLastPromotion),
            Median_YearsWithCurrManager = median(YearsWithCurrManager))
summary(MediansofLabtech)

#Medians for Manager
MediansofManager <- CaseStudy2_reorg %>% group_by(JobRole== "Manager") %>%
  summarise(Median_Age = median(Age),
            Median_DailyRate = median(DailyRate),
            Median_DistanceFromHome = median(DistanceFromHome),
            Median_HourlyRate = median(HourlyRate),
            Median_MonthlyIncome = median(MonthlyIncome),
            Median_MonthlyRate = median(MonthlyRate),
            Median_NumCompaniesWorked = median(NumCompaniesWorked),
            Median_PercentSalaryHike = median(PercentSalaryHike),
            Median_TotalWorkingYears = median(TotalWorkingYears),
            Median_TrainTimesLastYear = median(TrainingTimesLastYear),
            Median_YearsAtCompany = median(YearsAtCompany),
            Median_YearsINCurrentRole = median(YearsInCurrentRole),
            Median_YearsSinceLastPromotion = median(YearsSinceLastPromotion),
            Median_YearsWithCurrManager = median(YearsWithCurrManager))
summary(MediansofManager)

#Medians for Manufacturing Director
MediansofMD <- CaseStudy2_reorg %>% group_by(JobRole== "Manufacturing Director") %>%
  summarise(Median_Age = median(Age),
            Median_DailyRate = median(DailyRate),
            Median_DistanceFromHome = median(DistanceFromHome),
            Median_HourlyRate = median(HourlyRate),
            Median_MonthlyIncome = median(MonthlyIncome),
            Median_MonthlyRate = median(MonthlyRate),
            Median_NumCompaniesWorked = median(NumCompaniesWorked),
            Median_PercentSalaryHike = median(PercentSalaryHike),
            Median_TotalWorkingYears = median(TotalWorkingYears),
            Median_TrainTimesLastYear = median(TrainingTimesLastYear),
            Median_YearsAtCompany = median(YearsAtCompany),
            Median_YearsINCurrentRole = median(YearsInCurrentRole),
            Median_YearsSinceLastPromotion = median(YearsSinceLastPromotion),
            Median_YearsWithCurrManager = median(YearsWithCurrManager))
summary(MediansofMD)

#Medians for Research Director
MediansofRD <- CaseStudy2_reorg %>% group_by(JobRole== "Research Director") %>%
  summarise(Median_Age = median(Age),
            Median_DailyRate = median(DailyRate),
            Median_DistanceFromHome = median(DistanceFromHome),
            Median_HourlyRate = median(HourlyRate),
            Median_MonthlyIncome = median(MonthlyIncome),
            Median_MonthlyRate = median(MonthlyRate),
            Median_NumCompaniesWorked = median(NumCompaniesWorked),
            Median_PercentSalaryHike = median(PercentSalaryHike),
            Median_TotalWorkingYears = median(TotalWorkingYears),
            Median_TrainTimesLastYear = median(TrainingTimesLastYear),
            Median_YearsAtCompany = median(YearsAtCompany),
            Median_YearsINCurrentRole = median(YearsInCurrentRole),
            Median_YearsSinceLastPromotion = median(YearsSinceLastPromotion),
            Median_YearsWithCurrManager = median(YearsWithCurrManager))
summary(MediansofRD)


#Medians for Research Scientist
MediansofRS <- CaseStudy2_reorg %>% group_by(JobRole== "Research Scientist") %>%
  summarise(Median_Age = median(Age),
            Median_DailyRate = median(DailyRate),
            Median_DistanceFromHome = median(DistanceFromHome),
            Median_HourlyRate = median(HourlyRate),
            Median_MonthlyIncome = median(MonthlyIncome),
            Median_MonthlyRate = median(MonthlyRate),
            Median_NumCompaniesWorked = median(NumCompaniesWorked),
            Median_PercentSalaryHike = median(PercentSalaryHike),
            Median_TotalWorkingYears = median(TotalWorkingYears),
            Median_TrainTimesLastYear = median(TrainingTimesLastYear),
            Median_YearsAtCompany = median(YearsAtCompany),
            Median_YearsINCurrentRole = median(YearsInCurrentRole),
            Median_YearsSinceLastPromotion = median(YearsSinceLastPromotion),
            Median_YearsWithCurrManager = median(YearsWithCurrManager))
summary(MediansofRS)

#Medians for Sales Executive
MediansofSalesExecutive <- CaseStudy2_reorg %>% group_by(JobRole== "Sales Executive") %>%
  summarise(Median_Age = median(Age),
            Median_DailyRate = median(DailyRate),
            Median_DistanceFromHome = median(DistanceFromHome),
            Median_HourlyRate = median(HourlyRate),
            Median_MonthlyIncome = median(MonthlyIncome),
            Median_MonthlyRate = median(MonthlyRate),
            Median_NumCompaniesWorked = median(NumCompaniesWorked),
            Median_PercentSalaryHike = median(PercentSalaryHike),
            Median_TotalWorkingYears = median(TotalWorkingYears),
            Median_TrainTimesLastYear = median(TrainingTimesLastYear),
            Median_YearsAtCompany = median(YearsAtCompany),
            Median_YearsINCurrentRole = median(YearsInCurrentRole),
            Median_YearsSinceLastPromotion = median(YearsSinceLastPromotion),
            Median_YearsWithCurrManager = median(YearsWithCurrManager))
summary(MediansofSalesExecutive)


#Medians for Sales Representative
MediansofSalesRep <- CaseStudy2_reorg %>% group_by(JobRole== "Sales Representative") %>%
  summarise(Median_Age = median(Age),
            Median_DailyRate = median(DailyRate),
            Median_DistanceFromHome = median(DistanceFromHome),
            Median_HourlyRate = median(HourlyRate),
            Median_MonthlyIncome = median(MonthlyIncome),
            Median_MonthlyRate = median(MonthlyRate),
            Median_NumCompaniesWorked = median(NumCompaniesWorked),
            Median_PercentSalaryHike = median(PercentSalaryHike),
            Median_TotalWorkingYears = median(TotalWorkingYears),
            Median_TrainTimesLastYear = median(TrainingTimesLastYear),
            Median_YearsAtCompany = median(YearsAtCompany),
            Median_YearsINCurrentRole = median(YearsInCurrentRole),
            Median_YearsSinceLastPromotion = median(YearsSinceLastPromotion),
            Median_YearsWithCurrManager = median(YearsWithCurrManager))
summary(MediansofSalesRep)
```





```{r}

#Create function to calculate correlation and p-values

##Correlation matrix with p-values...see citation #1 below

cor.prob <- function (X, dfr = nrow(X) - 2) {

  R <- cor(X, use="pairwise.complete.obs")

  above <- row(R) < col(R)

  r2 <- R[above]^2

  Fstat <- r2 * dfr/(1 - r2)

  R[above] <- 1 - pf(Fstat, 1, dfr)

  R[row(R) == col(R)] <- NA

  R

}

 

# Create a function to dump the cor.prob output to a 4 column matrix with row/column indices, correlation, and p-value.

##See citation #2 below

flattenSquareMatrix <- function(m) {

  if( (class(m) != "matrix") | (nrow(m) != ncol(m))) stop("Must be a square matrix.")

  if(!identical(rownames(m), colnames(m))) stop("Row and column names must be equal.")

  ut <- upper.tri(m)

  data.frame(Explanatory_Var = rownames(m)[row(m)[ut]],

             Response_Var = rownames(m)[col(m)[ut]],

             cor=t(m)[ut],

             p=m[ut])

}

 

#Dummify all factor variables to use the correlation plot

Dummy <- dummyVars("~ .", data = CaseStudy2_reorg)

Case2Relab <- data.frame(predict(Dummy, newdata = CaseStudy2_reorg))



#Create correlation matrix with p-vales, then input into FlattenSquareMatrix to create a 4 column data frame
corMaster <- flattenSquareMatrix (cor.prob(Case2Relab))

#Check dataframe for 4 columns

print(head(corMaster,10))

 

#Order correlation in descending order with descending absolute correlation value

corMaster_reorg <- corMaster[order(-abs(corMaster$cor)),]

#Show 10 highest absolute correlation values

print(head(corMaster_reorg,10))


 

#filter the Response_Var for only Attrition.No outcomes

AttritionCor <- subset(corMaster_reorg, Response_Var == 'Attrition.No')

print(head(AttritionCor,20))


 

#filter the Response_Var for only Attrition.Yes outcomes

AttritionCor <- subset(corMaster_reorg, Response_Var == 'Attrition.Yes')

print(head(AttritionCor,10))


```

Citation #1 for Correlation matrix of p-values:

Title: correlation matricies: getting p-values?
Author: Bill Venables, Statistician, CMIS Environmetrics Project
Date: Tue Jan 4 06:05:39 CET 2000
Availability: <https://stat.ethz.ch/pipermail/r-help/2000-January/009758.html>

 

Citation #2 for FlattenSquareMatrix Function:

Title: Exploring correlations with R using cor.prob and chart.Correlation
Author: Stephen Turner
Date: August 27, 2012
Availability: https://gist.github.com/stephenturner/3492773


```{r}

#Create a model to classification model to predict attrition with sensitivty & specificity >= 60%



#Import dataset without Attrition
CaseStudyCompSet_NO_Attrition <- read.csv("C:/Users/Antonio Debouse/Desktop/Doing Data Science/GitHub/Case-Study-2/CaseStudy2CompSet No Attrition.csv")
str(CaseStudyCompSet_NO_Attrition)

#Change character variables to factors
CaseStudyCompSet_NO_Attrition[,names] <- lapply(CaseStudyCompSet_NO_Attrition[,names] , factor)


set.seed(12)
iterations = 100
masterAcc = matrix(nrow = iterations)
splitPerc = .7 #Training / Test split Percentage

for(j in 1:iterations)
{
  
  trainIndices = sample(1:dim(CaseStudy2_reorg)[1],round(splitPerc * dim(CaseStudy2_reorg)[1]))
  train = CaseStudy2_reorg[trainIndices,]
  test = CaseStudy2_reorg[-trainIndices,]
  
  model = naiveBayes(train[,c(12,15,8,9,7,11,25,28,21,17,30)],train$Attrition)
  table(predict(model,test[,c(12,15,8,9,7,11,25,28,21,17,30)]),test$Attrition)
  CM = confusionMatrix(table(predict(model,test[,c(12,15,8,9,7,11,25,28,21,17,30)]),test$Attrition))
  masterAcc[j] = CM$overall[1]
}

MeanAcc = colMeans(masterAcc)

MeanAcc
CM

#Input prediction into NO_Attrition dataset
AttritionPreds <- naiveBayes(CaseStudy2_reorg[,c(12,15,8,9,7,11,25,28,21,17,30)], CaseStudy2_reorg$Attrition)

CaseStudyCompSet_NO_Attrition$AttritionPreds = predict(AttritionPreds,CaseStudyCompSet_NO_Attrition)

#Create dataframe with ID and Attrition Predictions only
ID_AttritionPreds_Dataframe <- CaseStudyCompSet_NO_Attrition[,c("ID","AttritionPreds")]
ID_AttritionPreds_Dataframe

#Save data to CSV file
write.table(ID_AttritionPreds_Dataframe,"Case2PredictionsDebouse_Attrition.csv",sep = ",",row.names = FALSE)




#Knn with single k value

set.seed(7)
splitPerc = .7


trainIndices = sample(1:dim(CaseStudy2_reorg)[1],round(splitPerc * dim(CaseStudy2_reorg)[1]))
train = CaseStudy2_reorg[trainIndices,]
test = CaseStudy2_reorg[-trainIndices,]

classifications = knn(train[,c(17:30)],test[,c(17:30)],as.factor(train$Attrition), prob = TRUE, k = 10)
table(as.factor(test$Attrition),classifications)
CM = confusionMatrix(table(as.factor(test$Attrition),classifications))

CM

###Variable names used in knn model: "Age", "DailyRate", "DistanceFromHome", "HourlyRate", "MonthlyIncome", "MonthlyRate", "NumCompaniesWorked", "PercentSalaryHike", "TotalWorkingYears", "TrainingTimesLastYear", "YearsAtCompany", "YearsInCurrentRole", "YearsSinceLastPromotion","YearsWithCurrManager"

```






```{r}
#Create a multi-linear regression model to estimate MonthlyIncome with an RSME of <= 3000

##Create a correlation matrix plot

library(reshape2)
library(readxl)

#Import Dataset without Salary
CaseStudy2CompSet_NO_Salary <- read_excel("C:/Users/Antonio Debouse/Desktop/Doing Data Science/GitHub/Case-Study-2/CaseStudy2CompSet No Salary.xlsx")
view(CaseStudy2CompSet_NO_Salary)
str(CaseStudy2CompSet_NO_Salary)

#Change character variables to factors
CaseStudy2CompSet_NO_Salary[,names] <- lapply(CaseStudy2CompSet_NO_Salary[,names] , factor)

ggpairs(CaseStudy2_reorg[,17:31])

 

##MLR model to predict salary


SalaryPrediction_model = lm(MonthlyIncome ~ TotalWorkingYears + JobRole + Department + Gender+ BusinessTravel + DailyRate + JobLevel + PerformanceRating ,data = CaseStudy2_reorg)

summary(SalaryPrediction_model)
confint(SalaryPrediction_model)


#Cross Validation
set.seed(7)
TrainObs = sample(seq(1,dim(CaseStudy2_reorg)[1]),round(.75*dim(CaseStudy2_reorg)[1]),replace = FALSE)
SalaryTrain = CaseStudy2_reorg[TrainObs,]
SalaryTrain
SalaryTest = CaseStudy2_reorg[-TrainObs,]
SalaryTest
model1_fit = 
SalaryPrediction_model = lm(MonthlyIncome ~ TotalWorkingYears + JobRole + Department + Gender+ BusinessTravel + DailyRate + JobLevel + PerformanceRating ,data = SalaryTrain)
summary(model1_fit)




#Prediction Salary(MonthlyIncome) for the CaseStudy2 dataset with NO Salary 
CaseStudy2CompSet_NO_Salary$SalaryPreds = predict(SalaryPrediction_model,newdata = CaseStudy2CompSet_NO_Salary)


view(CaseStudy2CompSet_NO_Salary)

#Create dataframe with ID and Salary Predictions only
ID_SalaryPreds_Dataframe <- CaseStudy2CompSet_NO_Salary[,c("ID","SalaryPreds")]
ID_SalaryPreds_Dataframe

#Save data to CSV file
write.table(ID_SalaryPreds_Dataframe,"Case2PredictionsDebouseSalary.csv",sep = ",",row.names = FALSE)

```





